Covariate analysis of ROSMAP reprocessed counts (with CQN normalisation and covariates adjusted)

### 2017-04-14 code/Rmd/ROSMAP_REPROCESSED.Rmd
Minor modifications

knit2synapse::knitToFolderEntity(file = "./ROSMAP_REPROCESSED.Rmd",
                                 parentId = "syn8403743",
                                 entityName = 'ROSMAP Reprocessed RNASeq (Normalised)')
```

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn8403743';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of ROSMAP reprocessed RNASeq data';

@@

# Add in RIN^2 values
COVARIATES$RINcontinuous2 = COVARIATES$RINcontinuous^2
ContCovariates = c(ContCovariates, 'RINcontinuous2')
```
Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, fig.width=6, fig.height=6}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05)
draw(COVARIATES.CORRELATION$plot, heatmap_legend_side = 'left')
```
### Explore metadata
```{r data.explore, fig.width = 10, fig.height = 14}
# RIN
p = list()
p[[1]] = ggplot(COVARIATES, aes(x = Diagnosis, y = RINcontinuous)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + theme(legend.position = 'top')

@@

# Sex
# vcd::mosaic(~ msex + Diagnosis, data = COVARIATES)
```
### Filter genes
Remove genes that have less than 1 cpm counts in at least 50% of samples per cogdx. Also remove genes with missing gene length and percentage GC content
```{r cpmnormalisation}
genesToAnalyze = COVARIATES %>%
  rownameToFirstColumn('Sampleid') %>%
  dlply(.(Diagnosis), .fun = function(mtd, count){
  processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$Sampleid],
                                                   MIN_GENE_CPM=1, 
                                                   MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
  processed.counts$filteredExprMatrix$genes
}, COUNT)

@@

### Outlier Analysis
```{r outlier.analysis, cache = FALSE}
indToRemove = c('380_120503', '500_120515')

@@

METADATA$Diagnosis = 'OTHER'
METADATA$Diagnosis[METADATA$cogdx == 1 & METADATA$braaksc <= 3 & METADATA$ceradsc >= 3] = 'CONTROL'
METADATA$Diagnosis[METADATA$cogdx == 4 & METADATA$braaksc >= 4 & METADATA$ceradsc <= 2] = 'AD'

>>>>

knit2synapse::knitToFolderEntity(file = "./ROSMAP_REPROCESSED.Rmd",
                                 parentId = "syn8672910",
                                 entityName = 'ROSMAP Reprocessed RNASeq (Normalised)')
```

@@

```{r synapse.parameters, include=FALSE, cache=TRUE}
# Synapse parameters
parentId = 'syn8672910';
activityName = 'Covariate analysis';
activityDescription = 'Covariate analysis of ROSMAP reprocessed RNASeq data';

@@

# Add in RIN^2 values
COVARIATES$RINcontinuous2 = COVARIATES$RINcontinuous^2
ContCovariates = c(ContCovariates, 'RINcontinuous2')
```
Correlation/association between covariates at an FDR <= 0.1
```{r covariates.correlation, fig.width=9, fig.height=6}
COVARIATES.CORRELATION = getAssociationStatistics(COVARIATES, PVAL = 0.05)
draw(COVARIATES.CORRELATION$plot, heatmap_legend_side = 'left')
```
### Explore metadata
```{r data.explore, fig.width = 10, fig.height = 14}
# RIN
p = list()
p[[1]] = ggplot(COVARIATES, aes(x = Diagnosis, y = RINcontinuous)) + geom_boxplot()
p[[1]] = p[[1]] + ggtitle('RIN') + theme(legend.position = 'top')

@@

# Sex
# vcd::mosaic(~ msex + Diagnosis, data = COVARIATES)
```
### Filter genes
Remove genes that have less than 1 cpm counts in at least 50% of samples per Diagnosis. Also remove genes with missing gene length and percentage GC content
```{r cpmnormalisation}
genesToAnalyze = COVARIATES %>%
  rownameToFirstColumn('Sampleid') %>%
  dlply(.(Diagnosis), .fun = function(mtd, count){
  processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$Sampleid],
                                                   MIN_GENE_CPM=1, 
                                                   MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
  processed.counts$filteredExprMatrix$genes
}, COUNT)

@@

### Outlier Analysis
```{r outlier.analysis, cache = FALSE, fig.width = 10}
indToRemove = c('380_120503', '500_120515')

@@
